---
title: "Data Manipulation and Visualization"
author: "Instructors: Daniel Falster, Will Cornwell, Isaac Towers"
output: html_document
---

# Introduction to using Rmarkdown

- File extension `*.Rmd`
- Can be notebook or markdown
- Includes text and code -> make it good for note taking and reports 

Code chunks:

- run standard R code
- are embedded in a wrapper like this

```{r}

```

Text is written in [markdown](https://en.wikipedia.org/wiki/Markdown) --  a lightweight markup language with plain-text-formatting syntax.

- Headers: `#`
- Bold: `**bold**`
- links: `[text](link)`

The Rstudio interface makes it easy to work with both Rmarkdown (`*.Rmd`) and markdown files (`*.md`):

- easily run code chunks
- easily navigate
- use the document outline in the top corner to jump between headers

**Exercise**:  Create a new R markdown document (File -> New -> R Markdown, choose HTML). Add some text, then some code. Run the code chunks. Compile into an html file.   

# Introduction to movies dataset

For today's exercises we're using a data on movie production, from the ["Tidy Tuesday" collection](https://github.com/rfordatascience/tidytuesday). 

```{r}
library(readr)
library(ggplot2)
movies <- read_csv("data/movie_profit/movie_profit.csv")
```

What variables?
```{r}
names(movies)
```

An overview of the data

```{r}
#View(movies)
```

or summarise using the `skimr` package:

```{r}
skimr::skim(movies)
str(movies)
```

**Exercise**: With your partner, review the `README` for the data at `data/movie_profit/readme.md` to see the variables included. 


# What is the tidyverse?

* The [Tidyverse](http://tidyverse.org) is the name given to suite of R packages designed for seamless data analysis
* Designed to help you fall into a **"Pit of Success"**
* Tools are designed to work seamlessly together, for: 1) Turning data into tidy data, and 2) Plotting & analysing Tidy Data
* Not one but a collection packages
* Dataframes (tibbles) are the universal "tidy" input and output


Load (and install) individually or all together

```{r, eval = FALSE}
library(dplyr)
library(ggplot2)
```

# Reading data

Where possible, we recommend using 

- `csv` files (common separated files) to store data

```{r}
movies <- read.csv("data/movie_profit/movie_profit.csv")
```

Has a bunch of extra arguments: skip, check.names

**Exercises**: 

1. Review the help for `read.csv` and figure out what one of the above arguments does.
2. Use `read.csv` to load data file `data/plant_height/plantHeightSingleSpp.csv`
3. Use `read.csv` to load data file `data/sydney_beaches/temperature.csv`

```{r, exercise}
```

```{r, exercise}
```


When reading data, `read.csv` tries to guess column types. We can override this by specifying column types exactly.

Read everything as characters

```{r, exercise}
```

Also other variants

- `read.delim`: load file by specifying the delimiter

```{r, exercise}
```

# Data manipulation with `dplyr`

Motivation:

- Data is never organized in the way you want it 
- High % of project is data wrangling
- Many many many modern jobs are data wrangling

**Exercise**: Together with your partner, come up with 3 types of change you may need to make on a dataset before it is ready for analysis.




`dplyr` used verbs to describe the actions we want to take on the data

- `select`   -> subset columns 
- `filter`   -> subset rows 
- `arrange`  –> order rows
- `rename`   –> rename variables
- `mutate`   –> make new variables
- `summarise`–> summarise data
- `distinct` -> filter to each unique row

Examples:

To select certain variables:

```{r}
select(movies, genre)
```

To filter to particular rows:

```{r}
filter(movies, distributor == "Universal")
```

To sort by certain variables:

```{r}
arrange(movies, distributor, production_budget)
```

Create a new variable:

```{r}
mutate(movies, log_budget = log10(production_budget))
```

## Pipes

The pipe is a bit of magic. It's written by `%>%` (Shift-Command-M on Mac or Shift-Control-M on PC ). We can use "the pipe" [%>%](http://magrittr.tidyverse.org/reference/pipe.html) to connect expressions

* `%>%` is an **infix operator** -> expects commands on left & right
* Comes from the [magrittr](http://magrittr.tidyverse.org/reference/pipe.html) package
* `%>%` "pipes" the **output** of the last expression as the **first input** of the next expression
* If you use RStudio, you can type the pipe with Ctrl + Shift + M if you have a PC or Cmd + Shift + M if you have a Mac.

Examples:

```{r}
movies$distributor %>% unique()
```

```{r}
movies$distributor %>% unique() %>% length()
movies$distributor %>% n_distinct()
```

But you can control the input position of the next function with `.`:
```{r}
20 %>% seq(1, 4, length.out = .)
```


Tidyverse functions are written to work with pipes, i.e. most take the data as the first argument.

```{r}
filter(movies, distributor == "Universal")
```

is the same as

```{r}
movies %>% filter(distributor == "Universal")
```

This means we can use pipes to join data verbs to make a data sentence. 

```{r}
movies %>%
  filter(distributor == "Universal") %>%
  select(movie, mpaa_rating)
```

**Exercises:** Apply the `dplyr` package and your new data wrangling skills to the movies dataset to 


1. Create a subset of the dataset consisting of only movies distributed by `Walt Disney`
2. As above and only including variables `movie`, `worldwide_gross`, and `production_budget`
3. As above but with data sorted alphabetically by `movie`
4. As above but with an additional column return given by `worldwide_gross/production_budget`


```{r, exercise}
```

# Imagine your plot

https://www.data-to-viz.com

# Intro to data visualisation with `ggplot2`

See presentation in `slides` folder

**Examples from the slides:**

## An example

Let's start by building a ggplot from the ground up!

First all of all, what is my question? Although it might seem obvious, it can be really helpful to consider question you are trying to ask before starting to plot.In this case, my question: does the amount of money spent to create a movie vary by genre? For example, action movies might be more expensive compared to comedies because of all of the special effects required. 

Next, investigate the dataset to see if we have any data that will allow us to answer that question. 

```{r}
head(movies)
```


```{r}
movies %>%
  filter(production_budget > 1e7) %>% 
  ggplot(aes(x = genre, y = production_budget)) +
  geom_boxplot(outlier.colour = "red") +
  theme_classic() +
  xlab("Genre") -> p

ggsave("plot.png", p)
```

```{r}
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() 
```

```{r}
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() 
ggsave("figures/production_budget_boxplot.pdf")
```

## More data exploration plot excersizes

Use the ggplot cheat sheet!

Question 1: Is the distribution of domestic grosses across movies closer to a normal distribution or an log-normal distribution?

```{r, exercise}
```
Question 2:
Is there a relationship between production budget and domestic gross?

```{r, exercise}
```

Question 3:
Does the production budget and domestic gross relationship depend on genre?

```{r, exercise}
```
Question 4:
Can you make the last plot look better?

```{r, exercise}
```
