---
title: "Data Manipulation and Visualization"
author: "Instructors: Daniel Falster, Will Cornwell, Isaac Towers"
output: html_document
---

# Introduction to using Rmarkdown

- File extension `*.Rmd`
- Can be notebook or markdown
- Includes text and code -> make it good for note taking and reports 

Code chunks:

- run standard R code
- are embedded in a wrapper like this

```{r}

```

Text is written in [markdown](https://en.wikipedia.org/wiki/Markdown) --  a lightweight markup language with plain-text-formatting syntax.

- Headers: `#`
- Bold: `**bold**`
- links: `[text](link)`

The Rstudio interface makes it easy to work with both Rmarkdown (`*.Rmd`) and markdown files (`*.md`):

- easily run code chunks
- easily navigate
- use the document outline in the top corner to jump between headers

**Exercise**:  Create a new R markdown document (File -> New -> R Markdown, choose HTML). Add some text, then some code. Run the code chunks. Compile into an html file.   

# Introduction to movies dataset

For today's exercises we're using a data on movie production, from the ["Tidy Tuesday" collection](https://github.com/rfordatascience/tidytuesday). 

```{r}
library(tidyverse)
movies <- read_csv("data/movie_profit/movie_profit.csv")
```

What variables?
```{r}
names(movies)
```

An overview of the data

```{r}
#View(movies)
```

or summarise using the `skimr` package:

```{r}
skimr::skim(movies)
str(movies)
```

**Exercise**: With your partner, review the `README` for the data at `data/movie_profit/readme.md` to see the variables included. 


# What is the tidyverse?

* The [Tidyverse](http://tidyverse.org) is the name given to suite of R packages designed for seamless data analysis
* Designed to help you fall into a **"Pit of Success"**
* Tools are designed to work seamlessly together, for: 1) Turning data into tidy data, and 2) Plotting & analysing Tidy Data
* Not one but a collection packages
* Dataframes (tibbles) are the universal "tidy" input and output


Load (and install) individually or all together

```{r, eval = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
```
or
```{r, eval = FALSE}
library(tidyverse)
```

**Exercise**: Make sure you can load the `tidyverse`.

```{r}
library(tidyverse)
```



## What is a `tibble`? 

Basically, it's a better data frame.

Key features of a dataframe:

- table of data
- each column may have a different data type: character, double, integer, bool, string.

Tibbles have extra functionality:

* Prints a short summary at console
* Generates suitable warnings 
* Won't convert strings to factors
* No partial column-name matching
* Behaves consistently with `[` and `[[`
* For more, see [tibble.tidyverse.org](tibble.tidyverse.org)

```{r}
movies
```

# Reading data

Where possible, we recommend using 

- `csv` files (common separated files) to store data

```{r}
movies <- read.csv("data/movie_profit/movie_profit.csv")
```

Has a bunch of extra arguments: skip, check.names

**Exercises**: 

1. Review the help for `read.csv` and figure out what one of the above arguments does.
2. Use `read.csv` to load data file `data/plant_height/plantHeightSingleSpp.csv`
3. Use `read.csv` to load data file `data/sydney_beaches/temperature.csv`

```{r, exercise}
#2. Use `read.csv` to load data file `data/plant_height/plantHeightSingleSpp.csv`
read.csv("data/plant_height/plantHeightSingleSpp.csv")
```

```{r, exercise}
#3. Use `read.csv` to load data file `data/sydney_beaches/temperature.csv`
read.csv("data/sydney_beaches/temperature.csv")
```


When reading data, `read.csv` tries to guess column types. We can override this by specifying column types exactly.

Read everything as characters

```{r, exercise}
read.csv("data/movie_profit/movie_profit.csv", colClasses = "character")
```

Also other variants

- `read.delim`: load file by specifying the delimiter

```{r, exercise}
read.delim("data/movie_profit/movie_profit.csv", sep=",")
```

To load data from excel spreadsheets, try the `readxl` package.

```{r, exercise}

```


# Data manipulation with `dplyr`

Motivation:

- Data is never organized in the way you want it 
- High % of project is data wrangling
- Many many many modern jobs are data wrangling

**Exercise**: Together with your partner, come up with 3 types of change you may need to make on a dataset before it is ready for analysis.




`dplyr` used verbs to describe the actions we want to take on the data

- `select`   -> subset columns 
- `filter`   -> subset rows 
- `arrange`  –> order rows
- `rename`   –> rename variables
- `mutate`   –> make new variables
- `summarise`–> summarise data
- `distinct` -> filter to each unique row

Examples:

To select certain variables:

```{r}
select(movies, genre)
```

To filter to particular rows:

```{r}
filter(movies, distributor == "Universal")
```

To sort by certain variables:

```{r}
arrange(movies, distributor, production_budget)
```

Create a new variable:

```{r}
mutate(movies, log_budget = log10(production_budget))
```

## Pipes

The pipe is a bit of magic. It's written by `%>%` (Shift-Command-M on Mac or Shift-Control-M on PC ). We can use "the pipe" [%>%](http://magrittr.tidyverse.org/reference/pipe.html) to connect expressions

* `%>%` is an **infix operator** -> expects commands on left & right
* Comes from the [magrittr](http://magrittr.tidyverse.org/reference/pipe.html) package
* `%>%` "pipes" the **output** of the last expression as the **first input** of the next expression
* If you use RStudio, you can type the pipe with Ctrl + Shift + M if you have a PC or Cmd + Shift + M if you have a Mac.

Examples:

```{r}
movies$distributor %>% unique()
```

```{r}
movies$distributor %>% unique() %>% length()
movies$distributor %>% n_distinct()
```

But you can control the input position of the next function with `.`:
```{r}
20 %>% seq(1, 4, length.out = .)
```


Tidyverse functions are written to work with pipes, i.e. most take the data as the first argument.

```{r}
filter(movies, distributor == "Universal")
```

is the same as

```{r}
movies %>% filter(distributor == "Universal")
```

This means we can use pipes to join data verbs to make a data sentence. 

```{r}
movies %>%
  filter(distributor == "Universal") %>%
  select(movie, mpaa_rating)
```

**Exercises:** Apply the `dplyr` package and your new data wrangling skills to the movies dataset to 


1. Create a subset of the dataset consisting of only movies distributed by `Walt Disney`
2. As above and only including variables `movie`, `worldwide_gross`, and `production_budget`
3. As above but with data sorted alphabetically by `movie`
4. As above but with an additional column return given by `worldwide_gross/production_budget`


```{r, exercise}
# . Create a subset of the dataset consisting of only movies distributed by `Atlantic`
m1 <- movies %>% 
   filter(distributor=="Atlantic")
   
# 2. As above and only including variables `movie`, `worldwide_gross`, and `production_budget`
m2 <- m1 %>% 
   select(movie, worldwide_gross, production_budget)

# 3. As above but with data sorted alphabetically by `movie`
m3 <- m2 %>% 
   arrange(movie)

# 4.  As above but with an additional column return given by `worldwide_gross/production_budget`
m3 %>% 
   mutate(ratio = worldwide_gross/production_budget)

```

# Imagine your plot

https://www.data-to-viz.com

# Intro to data visualisation with `ggplot2`

See presentation in `slides` folder

**Examples from the slides:**

## An example

```{r}
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() 
```

## More data exploration plots
```{r}
#Density/Histograms
ggplot(data = movies, aes(x = log(domestic_gross))) + 
  geom_histogram(aes(y = ..density..)) +
  geom_density() 

#Scatter plots for exploring relationships
ggplot(data = movies, aes(x = production_budget, y = domestic_gross)) +   geom_point(alpha = 0.5)

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-sd(x)
   ymax <- m+sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}

#Violin plots with customised function
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_violin(trim = F) + 
  geom_point(colour = "navy", size = 1, alpha = 0.1) + 
  stat_summary(fun.data="data_summary", col = "black") 
```

