
## Colour schemes

### Continuous colour and discrete colour

### Colour-blindness

https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html

```{r}
library(viridis)
library(hexbin)

tibble(x = rnorm(10000), y = rnorm(10000)) %>%
  ggplot(aes(x = x, y = y)) +
  geom_hex() + coord_fixed() +
  scale_fill_viridis(option="D")+ 
  theme_bw()
```



```{r}
p <- data %>% 
  filter(year==1990) %>% 
  ggplot(aes(life_expectancy, gdp_per_capita)) + 
  geom_point(size=3, aes(colour = continent)) +
  scale_y_log10() +
  scale_color_viridis(discrete=TRUE) +
  theme_bw()

p
```

There are many more themes, e.g. from `ggsci` package

Nature Publishing Group
```{r}
library(ggsci)

p + scale_color_npg()
```

A different theme (American Academy of Science):
```{r}
p + scale_color_aaas()
```

## Font size

Change all the font sizes

```{r}
p + scale_color_aaas() +
theme(text = element_text(size=18)) 
```

Change just axis titles
```{r}
p + scale_color_aaas() +
theme(axis.title = element_text(size=18)) 
```



## Multiple planels

Load some extra packages

```{r}
library(janitor)
library(lubridate)
library(patchwork)
library(viridis)
```


```{r}
#Barplot production budget by Top 6 most common distributors
top_6 <- movies %>% tabyl(distributor) %>% arrange(-n) %>% 
  filter(n >= 240) %>% pull(distributor)

A <- movies %>% filter(distributor %in% top_6) %>% 
ggplot(aes(x = reorder(distributor, -production_budget, sum), y = production_budget, fill = distributor)) + 
  geom_col() + 
  ylab("Production budget (USD)") + 
  xlab("Distribution companies") + 
  scale_fill_manual(values = c("#A42820", "#5F5647", "#3F5151", "#4E2A1E", "#550307", "#0C1707")) + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "none")
A

#Side by side plot of domestic gross and worldwide gross of Top 6 most common distributors
distributor_long <- movies %>%
  filter(distributor %in% top_6) %>% 
  select(distributor, domestic_gross, worldwide_gross) %>% 
  pivot_longer(cols = 2:3, names_to = "type")

B <- ggplot(distributor_long, aes(x = reorder(distributor, -value, sum), y = value, fill = type)) + 
  geom_col(position = "dodge") +
  ylab("Gross income (USD)") + 
  xlab("Distribution companies") + 
  scale_fill_manual(values = c("#550307", "#0C1707"), name = "Type of income", labels = c("Domestic", "Worldwide")) + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10),
        axis.title.y = element_blank(),
        axis.text.y = element_blank())
B

#Plot release date by production budget
#Set release date as a date
movies <- movies %>% 
  mutate(release_date = mdy(release_date))

C <- ggplot(data = movies, aes(x = release_date, y = production_budget)) + 
  geom_point(alpha = 0.4, colour = "#0C1707") + 
  xlab("Release Date (Year)") + 
  ylab("Production budget (USD)") + 
  theme_bw() 
C

#Production cost and worldwide gross
D <- ggplot(data = movies, aes(x = production_budget, y = worldwide_gross)) + 
  geom_point(alpha = 0.4, colour = "#0C1707") + 
  geom_point(data = movies %>% filter(production_budget > 150000000 & worldwide_gross > 1000000000), colour = "#A42820") +
  geom_text(data = movies %>% filter(production_budget > 150000000 & worldwide_gross > 1000000000), aes(label = movie), size = 3, hjust = 1.1) + 
  geom_point(data = movies %>% filter(production_budget < 50000000 & worldwide_gross > 787000000), colour = "#5F5647") + 
  geom_text(data = movies %>% filter(production_budget  < 50000000 & worldwide_gross > 787000000), aes(label = movie), size = 3, hjust = -0.1) + 
  geom_point(data = movies %>% filter(production_budget < 100000000 & worldwide_gross > 1150000000), colour = "gold") + 
  geom_text(data = movies %>% filter(production_budget < 100000000 & worldwide_gross > 1150000000), aes(label = movie), size = 3, hjust = -0.2) + 
  ylab("Worldwide gross (USD)") + 
  xlab("Production budget (USD)") + 
  theme_bw() 
D

(A + B + C) / D
```


## Summarise and plot

```{r}
mpaa_rating_means <- movies %>% 
  filter(! is.na(mpaa_rating)) %>% 
  group_by(mpaa_rating) %>% 
  summarise(mean_worldwide_gross = mean(worldwide_gross),
            se_worldwide_gross = sd(worldwide_gross)/sqrt(length(worldwide_gross)),
            lower_worldwide_gross = mean_worldwide_gross - se_worldwide_gross,
            upper_worldwide_gross = mean_worldwide_gross + se_worldwide_gross) 

head(mpaa_rating_means)

ggplot(data = mpaa_rating_means, aes(x = mpaa_rating, y = mean_worldwide_gross)) + 
  geom_col() + 
  geom_errorbar(aes(ymin = lower_worldwide_gross, 
                    ymax = upper_worldwide_gross),
                width = 0.2)
  
```

## Customise "Accessorising"

```{r}
#Colour themes
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() + 
  theme_classic() 

ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() + 
  theme_bw() 

ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() + 
  theme_dark() 

ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() + 
  theme_void() 
```

```{r}
#Axis labels
ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() +
  ylab("Production budget (USD)") + 
  xlab("Genre") + 
  theme_bw() 

#Axis labels, colour, points and size/type face/font
ggplot(data = movies, aes(x = genre, y = production_budget, fill = genre)) + 
  geom_boxplot(outlier.colour = "red", size = 1.2) +
  ylab("Production budget (USD)") + 
  xlab("Genre") + 
  theme_bw() + 
  theme(axis.title.x = element_text(family = "Times", face = "bold.italic", size = 12),
        axis.title.y = element_text(family = "Times", face = "bold.italic", size = 12))
```

## Create your own theme - do it once and do it right

```{r}
thesis_theme <- theme_bw() + 
  theme(axis.title.x = element_text(family = "Calibri", face = "bold", size = 16),
        axis.title.y = element_text(family = "Calibri", face = "bold", size = 16),
        axis.text.x = element_text(family = "Calibri" , size = 12),
        axis.text.y = element_text(family = "Calibri", size = 12),
        strip.text = element_text(family = "Calibri", size = 12))

ggplot(data = movies, aes(x = genre, y = production_budget)) + 
  geom_boxplot() +
  ylab("Production budget (USD)") + 
  xlab("Genre") + 
  thesis_theme
```

### Facet wraps
```{r}
top6_plot <- movies %>% filter(distributor %in% top_6 & !is.na(mpaa_rating)) %>% 
  ggplot(data = ., aes(x = production_budget, y = domestic_gross)) +  
  geom_point() +
  theme_bw()

top6_plot + 
  facet_wrap(~distributor) 

top6_plot + 
  facet_grid(distributor ~ genre) 
```

### Scientific labels
https://robjhyndman.com/hyndsight/latex2exp/

```{r}
#Make some fake data
n <- 25
# the paramters
beta_0 <- 0.0005 # the true intercept
beta_1 <- 0.02 # the true slope
sigma <- 0.03 # the true standard deviation

temperature <- seq(20,60, by = 4)
metabolic_rate <- beta_0 + beta_1*temperature + rnorm(temperature, sd=sigma)
fake_data <- data_frame(temperature, metabolic_rate)

ggplot(fake_data, aes(x = temperature, y = metabolic_rate)) + 
  geom_point(size = 3) + 
  scale_y_log10() +
  ylab(expression(log~VCO_2~mL~min^{-1})) + 
  xlab(expression(Temperature^{ยบ}~C)) + 
  theme(axis.title.y = element_text(size = 30),
        axis.title.x = element_text(size = 30))
```

### Add a bit of flair: colour schemes
-Viridis, great for continuous data
-Rcolourbrewer Colour blind friendly palettes
-Wes Anderson palettes

```{r}
#Colour blind friendly
library(RColorBrewer)
display.brewer.all(colorblindFriendly = T)

ggplot(data = movies, aes(x = genre, y = production_budget, fill = genre)) + 
  geom_violin(trim = F, alpha = 0.7) + 
  geom_point(colour = "white", size = 1, alpha = 0.1) + 
  stat_summary(fun.data="data_summary", col = "black") + 
  scale_fill_brewer(palette = "Dark2") + 
  thesis_theme + 
  ylab("Production budget (USD)") + 
  xlab("Genres") +
  theme(legend.position = "none")

#Viridis
library(viridisLite)

ggplot(data = movies, aes(x = production_budget, y = worldwide_gross)) +
  geom_hex() + 
  scale_x_log10() + 
  scale_y_log10() + 
  scale_fill_viridis() + 
  ylab("Production budget (USD)") + 
  xlab("Worldwide gross (USD)") +
  thesis_theme

#Wes Andersson
library(wesanderson)

movies %>% filter(distributor %in% top_6) %>% 
ggplot(aes(x = reorder(distributor, -production_budget, sum), y = production_budget, fill = distributor)) + 
  geom_col() + 
  ylab("Production budget (USD)") + 
  xlab("Distribution companies") + 
  scale_fill_manual(values = wes_palette("IsleofDogs1")) + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "none")

movies %>% filter(distributor %in% top_6) %>% 
ggplot(aes(x = reorder(distributor, -production_budget, sum), y = production_budget, fill = distributor)) + 
  geom_col() + 
  ylab("Production budget (USD)") + 
  xlab("Distribution companies") + 
  scale_fill_manual(values = wes_palette("IsleofDogs1")) + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "none")
```

### Arranging multiple plots
https://patchwork.data-imaginist.com/
```{r}
#Install package
#install.packages('patchwork')
library(patchwork)

#Different layouts
A / D 

D / (A + C)

(A + B + C) / D

#Annotating
(A + B) / D + plot_annotation(tag_levels = 'A')
```


**Exercises**: Try to answer these questions below by making some relevant figures

1. Are there more cheap films?

 - Hint: create a histogram-plot with production_budget at the x-axis

```{r, exercise}
ggplot(movies, aes(x = production_budget)) +
  geom_histogram()
```
 
   
2. What genres make the highest return on investment?

 - Hints
   - calculate the ratio of worldwide gross profit and production budget
   - create a box-plot with x-axis representing the genre and y-axis representing the ratio
   - Rescale y-axis to log scale

```{r, exercise}
movies_ratio <- movies %>% mutate(ratio=worldwide_gross/production_budget)

ggplot(movies_ratio,aes(x = genre, y = ratio)) +
  geom_boxplot() +
   scale_y_log10()
```
   
3. Do expensive movies make more (gross) profit?

 - Hints
   - create a scatter plot with with the x-axis representing the production_budget and the y-axis representing the worldwide_gross profit
   - Rescale the x- and y-axis to log scale
   - add regression line with `stat_smooth()`

```{r, exercise}
# calculate the ratio of worldwide gross profit and production budget

movies_clean <- movies %>% filter(worldwide_gross!=0)

ggplot(movies_clean, aes(x = production_budget, y = worldwide_gross)) +
  geom_point(alpha=0.5) +
  stat_smooth(method = "lm", col = "red", se = FALSE)+
  scale_x_log10() + scale_y_log10()
```



# Customisation: 

**Exercise**: Brain storm how to improve this plot

## scales and axes

We can uses the scales package for customising labels and ticks

```{r}
library(scales) #  

# calculate profit
movies_profit <- movies %>% 
   mutate(profit = worldwide_gross - production_budget)


ggplot(movies_profit, aes(x = production_budget, y = profit, color=mpaa_rating)) + # set category/color as mpaa_rating
   geom_point(alpha = .2) + # alpha: set transparency to .2 or 20%
   stat_smooth(method = "lm", col="red", se=FALSE) + # add linear regression model
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) + # log transform x axis and display number in math notation
   scale_y_continuous(labels = comma) + # log transform y axis and display number in natural notation
   theme_bw() + # set theme
   xlab("Production budget in US$") + # change x label
   ylab("Profit in US$") + # change y label
   theme(legend.position = "none") # remove the legend (mpaa_rating)
   
```


## facet_wrap

Add a single line to split by 

```{r}

ggplot(movies_profit, aes(x = production_budget, y = profit, color=mpaa_rating)) + # set category/color as mpaa_rating
   geom_point(alpha = .2) + # alpha: set transparency to .2 or 20%
   stat_smooth(method = "lm", col="red", se=FALSE) + # add linear regression model
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) + # log transform x axis and display number in math notation
   scale_y_continuous(labels = comma) + # log transform y axis and display number in natural notation
   theme_bw() + # set theme
   xlab("Production budget in US$") + # change x label
   ylab("Profit in US$") + # change y label
   theme(legend.position = "none") + # remove the legend (mpaa_rating)
   facet_wrap(~genre)  # split figures by genre
  
```


## plotly

The package plotly can be used to make plots interactive.

First save the plot as an object and add a fiEld for label:

```{r}

p <- ggplot(movies_profit, aes(x = production_budget, y = profit, color=mpaa_rating, label = movie)) + # set category/color as mpaa_rating
   geom_point(alpha = .2) + # alpha: set transparency to .2 or 20%
   facet_wrap(~genre) + # split figures by genre
    scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10", scales::math_format(10^.x))) + # log transform x axis and display number in math notation
   scale_y_continuous(labels = comma) + # log transform y axis and display number in natural notation
   theme_bw() + # set theme
   xlab("Production budget in US$") + # change x label
   ylab("Profit in US$") # change y label
```

It is then very easy to turn this into an interactive plot -- you'll need to click on "Viewer" rather than "Plots". You can hover over the points to see the deatils of each.

```{r}
library(plotly)
ggplotly(p)
```



